Reproducible development environments with rix

### 2023-12-12 paper.qmd
more intro

@peng2011 introduced the idea of reproducibility being on a continuum: on one of
the ends of this continuum, we only have access to the paper describing the
studies, which is not reproducible at all. Then, if to this paper we make the
original source code of the analysis that was written to compute the results of
the study available, reproducibility is improved, albeit only by a little.
Adding the original data to this package improves reproducibility yet again. Finally,
if to all this we add what Roger Peng named the *linked and executable code and data*,
we reach the gold standard of full replication.

@@

- First, the version of [R]{.proglang} was stated

## Models and software {#sec-models}

The basic Poisson regression model for count data is a special case of the GLM framework @McCullaghNelder1989. It describes the dependence of a count response variable $y_i$ ($i = 1, \dots, n$) by assuming a Poisson distribution $y_i \sim \mathrm{Pois}(\mu_i)$. The dependence of the conditional mean $E[y_i \, | \, x_i] = \mu_i$ on the regressors $x_i$ is then specified via a log link and a linear predictor

$$
\log(\mu_i) \quad = \quad x_i^\top \beta,
$$ {#eq-mean}

where the regression coefficients $\beta$ are estimated by maximum likelihood (ML) using the iterative weighted least squares (IWLS) algorithm.

::: callout
TODO: Note that around the equation above there should be no spaces (avoided in the {{< latex >}} code by `%` lines) so that "normal" spacing is used and not a new paragraph started.
:::

>>>>

@peng2011 introduced the idea of reproducibility being on a continuum: on one of
the ends of this continuum, we only have access to the paper describing the
studies, which is not reproducible at all. Then, if in addition, to this paper
we make the original source code of the analysis that was written to compute the
results of the study available, reproducibility is improved, albeit only by a
little. Adding the original data improves reproducibility yet again. Finally, if
to all this we add what Roger Peng named the *linked and executable code and
data*, we reach the gold standard of full replication.

@@

- The version of [R]{.proglang} was stated;
- Packages and their versions were listed and saved into an `renv.lock` file;
- A `Makefile` was used to run the whole analysis and compile the paper;
- A `Dockerfile` was used to provide the complete computational environment, including the right version of [R]{.proglang} and run the whole analysis easily.

As we go down this list, we get closer to the gold standard of a perfectly reproducible
study.

However, reaching this gold standard is quite costly: one needs to learn a tool
to deal with package versions, then a build automation tool such as
[make]{.proglang} and ideally [Docker]{.proglang} should be added to the list.
[Docker]{.proglang} is a tool that makes it possible to run arbitrary code in a
completely controlled and isolated environment as a way to capture the complete
underlying system libraries.

### Prior art {#subsec-prior}
